### 2017-05-25 2017/NACO/paper.tex
Update Globalizer paper

(see, for instance, [12], [25], [38], [51]),

@@

various approaches to the dimensional reduction [38], [46], [48], [50], [51].

@@

reduction [46], [48], [50], [51].

@@

\section{Parallel Computations for Solving Global Optimization Problems}
\label{sec:parallel}
Let us consider the parallelization methods used widely in the theoretical and
practical application of parallel computing within the context of global optimization problems:
\begin{itemize}
  \item The distribution of the search domain \(D\) among the available computing units
  (data parallelization scheme). In the case of optimization problems, this approach is
  insufficient, since in organizing such computations the subdomain, which contain the
  sought global minimum, will be processed by only one processor, and, therefore,
  all of the remaining processors would perform the excess computations.
  \item The parallelization of the computational schemes for optimization algorithms
  (task parallelization scheme). This approach is also insufficient since the direct
  computational costs for executing optimization algorithms are relatively low
  (the majority of computations in the course of a global search are represented by
  calculations of the optimized function values due to the initial assumption of
  considerable computational costs for such calculations).
  \item The parallelization of the computations executed in order to obtain values
  for the optimized function. This approach can be applied since the most computationally
  costly part of the global search process will be parallelized. However, this method is not
  featured by the generality (the development of parallelization methods is to be performed
  every time from scratch while solving each particular optimization problem).
\end{itemize}
\par
Within the framework of information-statistical theory, a general approach to
parallelization computations when solving global optimization problems has been
proposed --- \textit{the parallelism of computations is provided by means of simultaneously
computing the values of the minimized function \(\varphi(y)\) at several different
points within the search domain \(D\)} – see, for example, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
This approach provides parallelizaion for the most costly part of computations in the global search process.
\par
The global optimization algorithms implemented in Globalizer will be described step-by-step below.
In Subsection \ref{subsec:corepar}, the core sequential multidimensional algorithm of global search (MAGS) will be presented.
In Subsection \ref{subsec:sharedpar}, a parallel generalization of the MAGS algorithm for parallel computations on
computer systems with shared memory will be described. In Subsection \ref{subsec:distribpar}, the scheme
for parallel computations by multiprocessor systems with distributed memory will be provided.

\subsection{Core Multidimensional Generalized Algorithm of Global Search}
\label{subsec:corepar}
The information-statistical theory of global search formulated in \cite{strongin1978}, \cite{strSergGO}
has served as a basis for the development of a large number of efficient multiextremal optimization methods --- see, for example, \cite{barkalovGergel2014}, \cite{Pizzuti}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin1994}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001} etc.
\par
Multidimensional Algorithm of Global Search (MAGS) established the basis for the
methods applied in Globalizer. The general computational scheme of MAGS can be
presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
\par
Let us introduce a simpler notation for the problem being solved on a computing node
\begin{equation}
\label{eq:oneDimFunc}
f(x) = \varphi(y(x)):x\in [0,1].
\end{equation}
\par
The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
Let us assume further \(k\), \(k>1\), iterations of a global search are to be completed.
The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
\par
\textit{Rule 1}. To renumber the points for the preceding trials by the lower indices in order of increasing value of coordinates
\begin{equation}
  \label{step1}
0=x_0<x_1<...<x_{k+1}=1.
\end{equation}
The points \(x_0\), \(x_k\) were introduced additionally for convenience in further explanation,
the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
\par
\textit{Rule 2}. To compute a current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
\begin{equation} \label{step2}
m=\max_{1\leq i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
M = \left\{
   \begin{array}{lr}
     rm, & m > 0,\\
     1, & m = 0,
   \end{array}
  \right.
\end{equation}
as the maximum of the relative differences of the minimized function \(f(x)\) from (\ref{eq:oneDimFunc})
on the set of executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}). Further, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\).
The constant \(r\), \(r>1\), is the \textit{parameter} for the algorithm.
\par
\textit{Rule 3}. For each interval \((x_{i-1},x_i),1\leq i\leq k+1\), compute the characteristics \(R(i)\) where
\[
R(i)=2\rho_i-4\frac{z_i}{M},\quad i=1,
\]
\begin{equation} \label{step3}
R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\frac{z_i+z_{i-1}}{M},\quad 1<i<k+1, \\
\end{equation}
\[
R(i)=2\rho_{i}-4\frac{z_{i-1}}{M},\quad i=k+1.
\]

\par
\textit{Rule 4}. To determine the interval with the maximum characteristic
\begin{equation} \label{step4}
R(t)=\max_{1\leq i \leq k+1}R(i).
\end{equation}
\par
\textit{Rule 5}. To execute a new trial (computation of the minimized function value \(f(x)\))
at the point \(x^{k+1}\) located within the interval with the maximum characteristic from (\ref{step4})
\begin{equation} \label{step5}
  x_{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{|z_{t}-z_{t-1}|}{M}\right]^N,\; \textrm{ if } 1<t<k+1,
\end{equation}
\[
  x_{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
\]

\par
The stop condition by which the trials are terminated, is defined by the inequality
\begin{equation}
  \label{eq:stop_1}
\rho_t<\varepsilon
\end{equation}
for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
accuracy of the problem solution. If the stop condition is not satisfied, the index \(k\) is incremented by unity,
and a new global search iteration is executed.
\par
In order to explain the algorithm presented above, let us note the following.
The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3})
could be interpreted as some measures of importance to the intervals with respect to
location of the global minimum point. Thus, the scheme for selecting the interval
for executing the next trial described in (\ref{step4}), (\ref{step5}) becomes more clear --- the point of every next
global search iteration is selected within the interval where the global minimum point can most likely be found.
\par
The conditions for the algorithm’s convergence described above were examined, for example, in \cite{strSergGO}.

\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}
Modern supercomputer computational systems consist of plenty of computer nodes,
which include several multicore processors. At that, random access memory is
shared for the computing nodes --- the content of any memory element can be read (written)
by any computer cores available at any arbitrary moment. In most cases, shared memory is
uniform --- the time characteristics for accessing memory are the same for all computer
cores and for all memory elements.
\par
The following speculations can form the basis for selecting parallel computation
organization methods. As was mentioned above when describing the MAGS algorithm,
the characteristics \(R(i),1\leq i\leq k+1\), calculated according to (\ref{step3})
can be interpreted as some measures of the importance of intervals with respect to the location of
the global minimum point. Following this understanding, each successive global search
iteration is executed within the most important interval with the maximum value of the
characteristic. So far, in this case it is obvious how to select the other intervals for
organizing simultaneous computations of the minimized function values at several different
points within the search domain as well --- these should be the intervals with the next
magnitudes of characteristics.
\par
The computational scheme of the Parallel Multidimensional Algorithm of Global Search for
computer systems with shared memory (PMAGS) was developed based the approach considered
above and is practically identical to the MAGS scheme --- the differences consist just in the
following set of rules --- see \cite{strSergGO}, \cite{stronginGergelBarkalovParGO}.
\par
\textit{Rule 4'}. To arrange the characteristics of the intervals obtained according to (\ref{step3}) in decreasing order
\begin{equation}
\label{step4par}
R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k})\geq R(t_{k+1})
\end{equation}
and to select \(p\) intervals with the indices \(t_j,1\leq j\leq p\), with
the highest values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
\par
\textit{Rule 5'}. To execute new trials (computations of values for the minimized
function \(f(x)\)) at the points \(x_{k+j},1\leq j\leq p\), located in the
intervals with the highest characteristics from (\ref{step4par})

\begin{equation} \label{step5par}
x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
\end{equation}
\[
x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\; \textrm{ if } t_j=1 \textrm{ or } t_j=k+1.
\]
\par
The stop condition for the algorithm (\ref{eq:stop}) should be checked for all
intervals from (\ref{step4par}), in which the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leq j\leq p.
\end{equation}
As before, if the stop condition is not fulfilled, the index \(k\) is
incremented by \(p\), and a new global search iteration is executed.
\par
The conditions for the developed parallel algorithm’s convergence and for the
non-redundancy of parallel computations have been examined in \cite{strSergGO}.
Thus, in particular, when the conditions for convergence are satisfied, the parallel
computations are non-redundant as compared to the sequential method when using up to
\(2^N\) processors/cores (where \(N\) is the dimensionality of the global optimization problem being solved).

@@

The next level of parallel computations in high-performance computer systems consists of
using several computational nodes. For that, each computational node has its own separate
memory and, in this case, the interaction between different nodes can be provided by
means of data transfer only via the computer communication network.

@@

To organize parallel computations for multiprocessor systems with distributed memory, using a set of mappings

@@

instead of applying a single Peano curve \(y(x)\) has been proposed in \cite{strongin1992}, \cite{strSergGO}.
In order to build the set \(Y_s(x)\), several different approaches can be applied.
Thus, for example, in \cite{strongin1992} a scheme has been applied by which each
mapping \(y_i(x)\) from \(Y_s(x)\) is built as the result of some shift along the main
diagonal of the hypercube \(D\). This way, the set of constructed Peano curves enables the
close prototypes \(x'\), \(x''\)  to be obtained from the interval \([0,1]\) for any close
multidimensional images \(y'\), \(y''\) from \(D\) differing in one coordinate,
only for some mapping \(y_k(x), 1\leq k\leq s\).

@@

Some other methods for constructing multiple mappings have been examined in \cite{stronginGergelBarkalovParGO}.

@@

The set of mappings \(Y_s(x)\) from (\ref{eq:maps}), for multidimensional problem (\ref{eq:task})
generates \(s\) subsidiary information-linked one-dimensional problems of the kind (\ref{eq:oneDimTask}):

@@

It is important to note that the family of one-dimensional problems
\varphi(y^l(x)),1 \leq l \leq s,
obtained as a result of the dimensional reduction is information-linked --- the function
values computed for any problem \(\varphi(y^l(x))\) from the family (\ref{eq:oneDimProblemSerie}) can be used for all
of the remaining problems of this family.

@@

The information compatibility of the problems from the family (\ref{eq:oneDimProblemSerie}) allows the parallel
computations to be organized in an obvious enough way. Thus, each particular problem can
be solved by a separate processor in the computing system; the exchange of search information
between the processors should be performed during the course of the computations.
As a result of using such a parallelization method, a unified approach for organizing parallel
computations for a multiprocessor computer systems with distributed, as well as shared
memory can be developed. This method consists of the following.

@@

  \item The family of reduced one-dimensional information-linked problems (\ref{eq:oneDimProblemSerie}) is
  distributed among the computational nodes of a multiprocessor system. One problem,
  as well as several others from within the family, can be allocated to each particular computational node.
  \item The Parallel Multidimensional Algorithm of Global Search (PMAGS, see Subsection \ref{subsec:sharedpar})
  is applied to solve the allocated problems from the family (\ref{eq:oneDimProblemSerie})
  at each computational node, supplemented by the following rules of information interaction.

@@

    \item Prior to beginning a new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y_l(x)),1\leq l\leq s\),
    the following should be performed:

@@

      \item compute the image \(y'\in D\) for the point \(x'\in [0, 1]\) according to the mapping \(y_l(x)\),
      \item determine  the prototypes \(x_i',1\leq i\leq s\), for each
      problem from the family (\ref{eq:oneDimProblemSerie}),
      \item send the prototypes \(x_i',1\leq i\leq s\) to all computational
      nodes employed in order to exclude the repeated selection of the intervals in
      which the prototypes fall, for using them to determine the points for new trials.
      To organize the data transfer, a queue can be formed at each computational
      node for sending the trial points and receiving the minimized function values at these points.

@@

    \item After completing any trial for any problem \(\varphi(y_l(x)),1\leq l\leq s\), at any point
    \(x'\in[0,1]\), it is necessary:

@@

      \item to determine all prototypes \(x_i',1\leq i\leq s\), for the point of the completed trial
      for each problem from the family (\ref{eq:oneDimProblemSerie}),
      \item to send the prototypes \(x_i',1\leq i\leq s\), and the result of the
      trial \(z'=\varphi(y_l(x'))\) to all computational nodes employed to include the
      obtained data into the search information processed within the rules of the parallel global search algorithm.

@@

    \item Prior to starting the next global search iteration, the algorithm should
    check the queue of the received data; if there are some data in the queue,
    they should be included in the search information.

@@

The possibility of asynchronous data transfer (the computational nodes process received
data only upon receipt) is a meaningful feature of such a scheme for organizing parallel computations.
In addition, there is no single managing node within this scheme. The number of
computational nodes can vary during the course of the global search, and excluding any
node does not result in the loss of the sought global minimum of the minimized multiextremal function.

@@

Global Search (GPMAGS) for high-performance computing systems, which may include many
multiprocessor (multicore) computational nodes with distributed memory as well as
accelerators for computations based on Intel Xeon Phi processors and on general purpose graphic processors.

@@

Additional information on organizing such parallel computation schemes is provided in \cite{gergelSidorov2015}.

\section{Globalizer System Architecture}
\label{sec:arch}
As previously mentioned in Section \ref{sec:intro}, the development of software systems
for global optimization is complex professional work. The optimization systems should
include advanced optimization methods, provide for the accumulation and efficient
processing of search information, include various forms of visualization for the results,
provide tools for interacting with users, etc. All the features pointed out above
become much more complicated when organizing parallel computations using high-performance
multiprocessor systems of varying architectures.
\par
The Globalizer considered in this paper expands the family of global optimization
software systems successively developed by the authors during the past several years.
One of the first developments was the SYMOP multiextremal optimization system \cite{gergel1993},
which has been successfully applied for solving many optimization problems. A special
place is occupied by the ExaMin system (see, for example, \cite{barkalovGergel2015}),
which was developed and used extensively to investigate the application of novel parallel
algorithms to solve global optimization problems using high-performance multiprocessor computing systems.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{pictures/globalizerScheme.eps}
    \caption{The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)}
    \label{fig:globalizerScheme}
\end{figure}

>>>>

(see, for instance, \cite{floudasPardalosGOState}, \cite{horstTuyGO}, \cite{pinterGO}, \cite{strSergGO}),

@@

various approaches to the dimensional reduction \cite{pinterGO}, \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.

@@

reduction \cite{sergeyevStronginLera2013}, \cite{strongin1978}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.

@@

\section{Parallel Computations for Solving Global Optimization Problems}
\label{sec:parallel}
Let us consider the parallelization approaches used widely in the theoretical and
practical applications of parallel computing within the context of global optimization problems:
  \item The distribution of the search domain \(D\) among the available computational units (data parallelization scheme). This approach is insufficient in the case of optimization problems since in such computations the subdomain, which contain the sought global minimum, will only be processed by one processor, and, therefore, all the remaining processors would perform the excess computations.
  \item The parallelization of the optimization algorithms (task parallelization scheme). This approach is also insufficient since the direct computational costs for executing the optimization algorithms are relatively low (the majority of computations are represented by the calculations of the optimized function values due to the initial assumption of considerable computational costs of such calculations).
  \item The parallelization of the computations executed in order to obtain the values of the optimized function. This approach is applicable since the most time-consuming part of the global optimization will be parallelized. However, this method is not featured by the generality (the development of parallelization methods is to be performed every time from the very beginning while solving each particular optimization problem).
Within the framework of the information-statistical theory, a general parallelization
approach for solving the global optimization problems has been proposed [\cite{stronginGergelBarkalovParGO}, \cite{strSergGO} --- \textit{the
parallel computations are provided by means of simultaneous computing the values of the
minimized function \(\varphi(y)\) at several different points within the search domain \(D\)}.
This approach provides the parallelization for the most time-consuming part of the optimization computations.
The global optimization algorithms implemented in Globalizer will be described step-by-step.
In Subsection \ref{subsec:corepar}, the core serial Multidimensional Algorithm of Global
Search (MAGS) is presented. In Subsection \ref{subsec:sharedpar}, a parallel generalization
of the MAGS algorithm for the parallel computations on the multiprocessor computational
nodes with shared memory is described. In Subsection \ref{subsec:distribpar}, the scheme
for the parallel computations on the high-performance computational systems with distributed memory is given.
\subsection{Core Multidimensional Generalized Algorithm of Global Search}
\label{subsec:corepar}
The information-statistical theory of global optimization formulated in \cite{strongin1978}, \cite{strSergGO} has
served as a basis for the development of a large number of efficient multiextremal optimization
methods \cite{barkalovGergel2014}, \cite{Pizzuti}, \cite{gergel1996}, \cite{gergel1997}, \cite{grishaginStrongin1984}, \cite{sergeyev1995}, \cite{sergeyev1999}, \cite{sergeyevGrishagin2001}, \cite{sergeyevStronginLera2013}, \cite{Famularo2001}.
The optimization methods applied in Globalizer are based on the MAGS method, which can be presented as follows --- see \cite{strongin1978}, \cite{strSergGO}.
Let us introduce a simpler notation for the optimization problem being solved
\label{eq:oneDimFunc}
f(x) = \varphi(y(x)):x\in [0,1].
The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
Then, let us suppose that \(k\), \(k\ge 1\), optimization iterations have been completed already.
The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
\textit{Rule 1}. Renumber the points of the preceding trials by the lower indices in order of increasing value of coordinates
  \label{step1}
0=x_0<x_1<...<x_{k+1}=1.
the points \(x_0\), \(x_k\) were introduced additionally for the convenience of further
explanation, the values of the minimized function \(z_0\), \(z_k\) at these points are undefined.
\textit{Rule 2}. Compute current estimate of the H{\"o}lder constant \(H\) from (\ref{eq:holder})
\begin{equation} \label{step2}
m=\max_{1\leq i\leq k}\frac{|z_i-z_{i-1}|}{\rho_i}, \;
M = \left\{
   \begin{array}{lr}
     rm, & m > 0,\\
     1, & m = 0,
  \right.
as the maximum of the relative differences of the minimized function values on the
set of previously executed trial points \(x_i,1\leq i\leq k\) from (\ref{step1}).
Hereafter, \(\rho_i=(x_i-x_{i-1})^\frac{1}{N},1\leq i\leq k+1\). The
constant \(r\), \(r>1\), is the reliability parameter of the algorithm.
\textit{Rule 3}. Compute the characteristics \(R(i)\) for each interval \((x_{i-1},x_i),1\leq i\leq k+1\), where
R(i)=2\rho_i-4\frac{z_i}{M},\quad i=1,
\begin{equation} \label{step3}
R(i)=\rho_i+\frac{(z_i-z_{i-1})^2}{M^2\rho_i}-2\frac{z_i+z_{i-1}}{M},\quad 1<i<k+1, \\
R(i)=2\rho_{i}-4\frac{z_{i-1}}{M},\quad i=k+1.

\par
\textit{Rule 4}. Determine the interval with the maximum characteristic
\begin{equation} \label{step4}
R(t)=\max_{1\leq i \leq k+1}R(i).
\end{equation}
\par
\textit{Rule 5}. Execute a new trial at the point \(x^{k+1}\) located within the interval
with the maximum characteristic from (\ref{step4})
\begin{equation} \label{step5}
  x_{k+1}=\frac{x_t+x_{t-1}}{2}-\mathrm{sign}(z_{t}-z_{t-1})\frac{1}{2r}\left[\frac{|z_{t}-z_{t-1}|}{M}\right]^N,\; \textrm{ if } 1<t<k+1,
\end{equation}
\[
  x_{k+1}=\frac{x_t+x_{t-1}}{2},\; \textrm{ if } t=1 \textrm{ or } t=k+1.
\]

\par
The stop condition, which terminated the trials, is defined by the inequality
\begin{equation}
  \label{eq:stop_1}
\rho_t<\varepsilon
\end{equation}
for the interval with the maximum characteristic from (\ref{step4}) and \(\varepsilon >0\) is the predefined
accuracy of the optimization problem solution. If the stop condition is not satisfied,
the index \(k\) is incremented by 1, and new global optimization iteration is executed.
\par
In order to explain the algorithm presented above, let us note the following.
The characteristics \(R(i), 1\leq i\leq k+1\), calculated according to (\ref{step3}) could
be interpreted as some measures of importance of the intervals with respect to the
expected location of the global minimum point. Thus, the rules (\ref{step4}) and (\ref{step5}) for selecting
the interval for executing the next trial become more clear --- the point of every next
global optimization iteration is selected within the interval, where the global minimum
point can be found most likely.
\par
The convergence conditions for the described algorithm are given, for example, in \cite{strSergGO}.

\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}
Modern supercomputer computational systems consist of plenty of computational nodes,
which include several multicore processors. In this case, random access memory is
shared among the computational nodes --- the content of any memory location can be read
(written) by any computational core at any arbitrary moment. In most cases, shared memory is
uniform --- the time characteristics for accessing the memory are the same for all
computational cores and for all memory locations.
\par
The following approach can be applied to provide the parallel computations with the MAGS method.
As it was mentioned above, the characteristics \(R(i),1\leq i\leq k+1\) calculated
according to (\ref{step3}) can be interpreted as some measures of importance of the
intervals with respect to the expected location of the global minimum point.
Following this understanding, every next global optimization iteration is executed within
the most important interval with the maximum characteristic. As a result, it becomes
evident how to select the other intervals for simultaneous computations of the minimized
function values at several different points within the search domain --- these ones
should be the intervals with the next magnitudes of characteristics.
\par
The computational scheme of the Parallel Multidimensional Algorithm of Global Search (PMAGS)
for the computational systems with shared memory can be developed based on the approach
described above. This scheme is almost identical to the MAGS one --- the differences
consist just in the following set of rules.
\par
\textit{Rule 4'}. Arrange the characteristics of the intervals computed according to (\ref{step3}) in the decreasing order
\begin{equation}
\label{step4par}
R(t_1)\geq R(t_2)\geq \dots \geq R(t_{k})\geq R(t_{k+1})
\end{equation}
and select \(p\) intervals with the indices \(t_j,1\leq j\leq p\), with the maximum
values of characteristics (\(p\) is the number of processors/cores used for the parallel computations).
\par
\textit{Rule 5'}. Execute new trials at the points \(x_{k+j},1\leq j\leq p\) located
in the intervals with the maximum characteristics from (\ref{step4par})

\begin{equation} \label{step5par}
x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2}-\mathrm{sign}(z_{t_j}-z_{t_j-1})\frac{1}{2r}\left[\frac{|z_{t_j}-z_{t_j-1}|}{M}\right]^N,\; \textrm{ if } 1<t_j<k+1,
\end{equation}
\[
x_{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2},\; \textrm{ if } t_j=1 \textrm{ or } t_j=k+1.
\]
\par
The stop condition (\ref{eq:stop_1}) should be checked for all intervals from (\ref{step4par}),
where the scheduled trials are executed
\begin{equation}
  \label{eq:stop}
\rho_{t_j}<\varepsilon,1\leq j\leq p.
\end{equation}
If the stop condition is satisfied at least for one interval \(t_j, 1\le j\le p\), the calculations
are terminated. Otherwise, the index \(k\) is incremented by \(p\), and new global optimization iteration is executed.
\par
The convergence conditions for the developed parallel algorithm are considered
in \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}. Thus, in particular, when the convergence conditions are
satisfied, the parallel computations are non-redundant as compared to the serial method
when using up to \(2^N\) processors/cores (\(N\) is the dimensionality of the global optimization problem).

@@

The next level of parallel computations on the high-performance computational systems
consists in using several computational nodes. In this case, each computational node has
its own separate memory and the interaction between different nodes can be provided by
means of data transfer via the computer communication network.

@@

To provide parallel computations on high-performance computational systems with the distributed memory, it is proposed to use a set of mappings

@@

instead of applying a single Peano curve \(y(x)\) --- see, for instance,
\cite{strongin1992}, \cite{stronginGergelBarkalovParGO}, \cite{strSergGO}.
In order to construct the set \(Y_s(x)\), several different approaches can be applied.
Thus, for example, in \cite{strongin1992} each mapping \(y^i(x)\) from \(Y_s(x)\) is constructed
as the result of some shift along the main diagonal of the hypercube \(D\). This way,
the set of the constructed Peano curves enables the close prototypes \(x'\), \(x''\)
to be obtained from the interval \([0, 1]\) for any close multidimensional images
\(y'\), \(y''\) from \(D\) differing in one coordinate only.

@@

Some other methods for constructing the multiple mappings were considered in \cite{stronginGergelBarkalovParGO}.

@@

The set of mappings \(Y_s(x)\) from (\ref{eq:maps}) generates \(s\) information-linked one-dimensional
problems of the kind (\ref{eq:oneDimTask}):

@@

It is important to note that the family of the one-dimensional problems
\(\varphi(y^l(x)),1 \leq l \leq s\) obtained as a result of the dimensionality
reduction is an information-linked one --- the function values computed for any
problem \(\varphi(y^l(x))\) from the family (\ref{eq:oneDimProblemSerie}) can be used for all rest problems of this family.

@@

The information compatibility of the optimization problems from the family (\ref{eq:oneDimProblemSerie})
allows executing the parallel computations in the following way. Each particular
problem can be solved by a separate processor of the computational system; the exchange
of search information between the processors should be performed during the computations.
As a result, a unified approach for the parallel computations on the high-performance
computational systems with distributed and shared memory can be developed. This method consists in the following.

@@

  \item The family of the reduced one-dimensional information-linked optimization
  problems (\ref{eq:oneDimProblemSerie}) is distributed among the computational nodes of the multi-node computational system.
  \item Each optimization problem from the family (\ref{eq:oneDimProblemSerie}) is solved on the
  corresponding computational node using the PMAGS method (described above in Subsection \ref{subsec:sharedpar})
  supplemented by the following rules of the information interaction.

@@

    \item Prior to beginning new trial at any point \(x'\in [0,1]\) for any problem \(\varphi(y^l(x)),1\leq l\leq s\), the following should be performed:

@@

      \item compute the image \(y'\in D\) for the point \(x'\in [0, 1]\) according to the mapping \(y^l(x)\),
      \item determine  the preimages\footnote{All preimages have the same
      image \(y'\); the preimages are calculated by the inverse mappings \((y^i)^{-1}\), \(1\le i\le s\);
      the number of the preimages coincides with the number of the mappings.}
      \begin{displaymath}
        x_i':y'=y^i(x_i'),1\le i\le s,
      \end{displaymath}
      in accordance with the set of mappings \(y^l(x), 1\le l\le s\) for each problem from the family (\ref{eq:oneDimProblemSerie}),
      \item send the prototypes \(x_i',1\leq i\leq s\) to all computational nodes in order
      to exclude the repeated selection of the intervals, which the prototypes fall in,
      for using these ones to determine the points of new trials. To perform the data transfer,
      a queue can be created at each computational node for sending the trial points and
      receiving the minimized function values at these points.

@@

    \item After completing any trial for any problem \(\varphi(y_l(x)),1\leq l\leq s\),
    at any point \(x'\in[0,1]\), it is necessary:

@@

      \item to determine all prototypes \(x_i',1\leq i\leq s\) for the point of the
      completed trial for each problem from the family (\ref{eq:oneDimProblemSerie}),
      \item to send the prototypes \(x_i',1\leq i\leq s\) and the result of the trial
      \(z'=\varphi(y_l(x'))\) to all computational nodes to include the obtained data into
      the search information processed according to the rules of the parallel global optimization algorithm.

@@

    \item Prior to starting the next global optimization iteration, the algorithm
    should check the queue of the received data; if there are any data in the queue,
    these should be included into the search information. The MAGS algorithm uses these
    accumulated data for the adaptive computing of the optimization iteration points (see Section \ref{subsec:corepar}).

@@

The possibility of asynchronous data transfer (the computational nodes process the received
data upon receipt only) is an important feature of the proposed parallel computation scheme.
In addition, there is no any selected managing node within this scheme. The number of
computational nodes can vary during the global optimization, and excluding any node does not result
in the loss of the sought global minimum of the minimized multiextremal function.

@@

Global Search (GPMAGS) for high-performance computational systems, which may include many
multiprocessor (multicore) computational nodes with distributed memory as well as the
computational accelerators based on Intel Xeon Phi processors and on the general purpose graphic processors (GPUs).

@@

Additional information on such parallel computation schemes can be found in \cite{gergelSidorov2015}.

\section{Globalizer System Architecture}
\label{sec:arch}
Globalizer expands the family of the global optimization software systems developed
by the authors during the past several years. One of the first developments was the
SYMOP multiextremal optimization system \cite{gergel1993}, which has been applied for solving
many optimization problems. The ExaMin system \cite{barkalovGergel2015}, [3], [4], [16] was developed
and used to investigate various parallel algorithms for solving the global
optimization problems on the high-performance computational systems.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{pictures/globalizerScheme.eps}
    \caption{The architecture of the Globalizer system}
    \label{fig:globalizerScheme}
\end{figure}

\par
The architecture of the Globalizer system is presented in Figure \ref{fig:globalizerScheme}.
The system components are:
\begin{itemize}
  \item Block 0 consists of the procedures for computing the function values
  (criteria and constraints) of the optimization problem being solved; it is an external
  block with a predefined interface with respect to the Globalizer system and is provided by the user.
  \item Blocks 1-4 form the optimization subsystem and solve the global optimization
  problems (Block 1), nonlinear programming ones (Block 2), multicriterial optimization problems
  (Block 3), and general decision making problems (Block 4). The interaction between these
  components is organized in a hierarchical order --- the decision making problems are
  solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, and so on.
  \item Block 5 is a subsystem for accumulating and processing the search information; it is one of the main subsystems – the size of search information may appear to be quite large and the efficiency of the global optimization methods depends on how completely the available search data are utilized.
  \item Block 6 contains the dimensionality reduction procedures based on the Peano space-filling curves; the reduced (one-dimensional) optimization problems are solved by the optimization blocks. This block also provides the interaction between the optimization blocks and the multidimensional optimization problem being solved.
  \item Block 7 provides tuning the parallel computation schemes in the Globalizer system subject to the employed computational system architecture (the number of the computational cores, the availability of shared and/or distributed memory, the availability of computational accelerators, etc.) and the applied global optimization methods.
  \item Block 8 is responsible for managing the parallel processes when performing the global optimization (determining the optimal configuration of parallel processes, distributing the processes between the computational unit, balancing the computational loads, etc.).
  \item Block 9 is a managing subsystem, which controls the parallel computations when solving the global optimization problems.
  \item Block 10 provides the dialog interaction with the users for stating the optimization problem, adjusting the system parameters (if necessary), and visualizing and presenting the global optimization results.
  \item Block 11 is a set of tools for visualizing and presenting the global optimization results; the availability of the tools for visual demonstration of the computational results enables the user to perform an efficient control over the global optimization process.
\end{itemize}
\par
The Globalizer system comes in two different versions:
\begin{itemize}
  \item Globalizer Pro is a research version for making various large-scale numerical experiments on the high-performance computational systems to investigate the new parallel global optimization methods being developed.
  \item Globalizer Lite is a limited version for solving the global optimization problems with a medium complexity level, and is accessible free of charge.
\end{itemize}

\section{Results of Numerical Experiments}
\label{sec:experiments}
The numerical experiments were carried out using the Lobachevsky supercomputer at
the State University of Nizhni Novgorod (the operation system – CentOS 6.4, the
supercomputer management system – SLURM). Each supercomputer node had 2 Intel Sandy Bridge
E5-2660 processors, 2.2 GHz, with 64 GB RAM. Each processor had 8 cores (i. e. total
16 CPU cores were available at each node). In order to generate the executable program code, the Intel C++ 14.0.2 compiler was used.
\par
A number of computational nodes had Intel Xeon Phi 5110P processors, each containing 60 cores
(total 240 threads). In addition, the supercomputer included the computational nodes equipped
with two NVIDIA Kepler K20X GPUs, each providing 14 stream multiprocessors (total 2688 CUDA-cores). In order to generate the executable code, the CUDA Toolkit 6.0 was used.
\par
In the performed experiments the optimization problems generated by the GKLS-generator
\cite{gavianoKvasovLeraSergeev2003} were used. This generator allows obtaining the multiextremal optimization
problems with \textit{a priori} known properties: the number of local minima, the sizes of the attractors of these ones, the global minimum point, the function value at this point, etc. In order to simulate the computational costs inherent to the applied optimization problems, in all experiments computing the objective function was intentionally complicated by the auxiliary computations not altering the form of the function and the position of its minima.
\par
The numerical experiments were carried out first using the ExaMin system and then were reproduced using the Globalizer system. In all the executed experiments, 100 global optimization problems generated randomly by the GKLS-generator were solved. The results of experiments were averaged over the number of the solved optimization problems.

\subsection{Comparison with over methods}
The results of numerical comparison of three sequential algorithms --- DIRECT
\cite{Jones}, DIRECT\textit{l} \cite{Gablonsky}, and MAGS from Section \ref{subsec:corepar} --- are presented below.
